A semantics of core AUTOSAR

### 2016-05-16 paper/paper.tex
Paper macro fixes

The process terms of our calculus are as follows:
\begin{itemize}
\item $\runnable{\Tr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\rinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\excl{\Tx{X}{I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\irv{\Ts{S}{I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\delem{\Tep{E}{P}{I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\qelem{\Tep{E}{P}{I}}{\Vs}{TODO}$ \newline
Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$.
\item $\opres{\Top{O}{P}{I}}{\Vs}$ \newline
Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
\item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
\end{itemize}

In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls. Meta-variables $N$, $B$ and $T$ range over natural numbers, boolean values and (floating point) time values, respectively. The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise. An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\serving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls). Formally,
\[
\begin{array}{rcll}
	V	& \in		& \text{Values} \\
	N	& \in		& \text{Natural numbers} \\
	B	& \in		& \text{Boolean values} \\
	T	& \in		& \text{Time values} \\
	C	& ::=		& \Top{O}{P}{I} \sep \Tvoid \\
	\Act	& ::=		& \Tidle \sep \Tpending \sep \serving{\Cs}{\Vs} \\
\end{array}
\]
We write $\Vs$ for a sequence of values, $\Cs$ for a sequence of clients, etc. Sequences may be empty, which is written $\epsilon$. Left (right) concatenation of an element with a sequence is written $\Tcons{\V{V}}{\V{Vs}}$ ($\Tsnoc{\V{Vs}}{\V{V}}$).

@@

Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
\[
\begin{array}{rcll}
	P,Q 	& ::= &
		\runnable{A}{K}{T}{\Act}{N}	\\
		& | &
		\rinst{A}{C}{\Xs}{\Code}	\\
		& | &
		\excl{A}{B}			\\
		& | &
		\irv{A}{V}				\\
		& | &
		\delem{A}{B}{V}			\\
		& | &
		\qelem{A}{\Vs}{TODO}			\\
		& | &
		\opres{A}{\Vs}			\\
		& | &
		\timer{A}{T}		\\
		& | &
		P|Q					\\
\end{array}
\]

@@

\item For each runnable $R$ of each component $I$, a term
\[
\begin{array}{c}
	\runnable{\Tr{R}{I}}{TODO}{0.0}{\Act}{0}
\end{array}
\]
where $\Act$ is $\serving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
% SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable

\item For each exclusive area $X$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\excl{\Tx{X}{I}}{\Tfalse}
\end{array}
\]

\item For each inter-runnable variable $S$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\irv{\Ts{S}{I}}{V}
\end{array}
\]
where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\Tvoid$.

\item For each data element $E$ of each required port $P$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\qelem{\Tep{E}{P}{I}}{\epsilon}{TODO}
\end{array}
\]
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
\[
\begin{array}{c}
	\delem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
\end{array}
\]

\item For each operation $O$ of each required port $P$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\opres{\Tep{O}{P}{I}}{\epsilon}
\end{array}
\]

\item For each \emph{TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\timer{\Tr{R}{I}}{0.0}
\end{array}
\]
% where $T$ is the period of the event.
%TODO: Does this mean all timed runnables will want to act _immediately_ (at the same time)? No init delay?

>>>>

The process terms of our calculus are as follows:
\begin{itemize}
\item $\Trunnable{\Tr{R}{I}}{K}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$ with continuation $K$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\Trinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\Texcl{\Tx{X}{I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\Tirv{\Ts{S}{I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\Tdelem{\Tep{E}{P}{I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\Tqelem{\Tep{E}{P}{I}}{N}{\Vs}$ \newline
Queued sender-receiver data element $E$ in port $P$ of component $I$, currently holding value sequence $\Vs$. TODO Maximum queue capacity? $N$.
\item $\Topres{\Top{O}{P}{I}}{\Vs}$ \newline
Client-server operation $O$ in port $P$ of component $I$, currently holding result value sequence $\Vs$.
\item $\Ttimer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
\end{itemize}

In the untyped setting of this report, a value $V$ can be any data item computed and communicated by an AUTOSAR system, including the unit value $\Tvoid$ and the error codes that might be returned from RTE calls. Meta-variables $N$, $B$ and $T$ range over natural numbers, boolean values and (floating point) time values, respectively. The client $C$ of a runnable instance is the address of the invoking port operation (if the runnable was invoked by a client-server call), or $\Tvoid$ otherwise. An activation state $\Act$ either toggles between $\Tidle$ and $\Tpending$, or it has the special shape $\Tserving{\Cs}{\Vs}$ (in the case of a runnable triggered by client-server calls). Formally,
\[
\begin{array}{rcll}
	V	& \in		& \text{Values} \\
	N	& \in		& \text{Natural numbers} \\
	B	& \in		& \text{Boolean values} \\
	T	& \in		& \text{Time values} \\
	C	& ::=		& \Top{O}{P}{I} \sep \Tvoid \\
	\Act	& ::=		& \Tidle \sep \Tpending \sep \Tserving{\Cs}{\Vs} \\
\end{array}
\]
We write $\Vs$ for a sequence of values, $\Cs$ for a sequence of clients, etc. Sequences may be empty, which is written $\epsilon$. Left (right) concatenation of an element with a sequence is written $\Tcons{\V{V}}{\V{Vs}}$ ($\Tsnoc{\V{Vs}}{\V{V}}$).

@@

Together with an operator for composing process terms in parallel, the process grammar for our AUTOSAR calculus becomes
\[
\begin{array}{rcll}
	P,Q 	& ::= &
		\Trunnable{A}{K}{T}{\Act}{N}	\\
		& | &
		\Trinst{A}{C}{\Xs}{\Code}	\\
		& | &
		\Texcl{A}{B}			\\
		& | &
		\Tirv{A}{V}				\\
		& | &
		\Tdelem{A}{B}{V}			\\
		& | &
		\Tqelem{A}{N}{\Vs}			\\
		& | &
		\Topres{A}{\Vs}			\\
		& | &
		\Ttimer{A}{T}		\\
		& | &
		P|Q					\\
\end{array}
\]

@@

\item For each runnable $R$ of each component $I$ with continuation (code) $K$, a term
\[
\begin{array}{c}
	\Trunnable{\Tr{R}{I}}{K}{0.0}{\Act}{0}
\end{array}
\]
where $\Act$ is $\Tserving{\epsilon}{\epsilon}$ if $R$ is triggered by an \emph{OperationInvokedEvent}; else $\Act$ is $\Tpending$ if $R$ is triggered by an \emph{InitEvent}; otherwise $\Act$ is $\Tidle$.
% SWS_Rte_03524 (SWS_Rte_03526?) forbids mixing OperationInvokedEvents with other events for the same runnable

\item For each exclusive area $X$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\Texcl{\Tx{X}{I}}{\Tfalse}
\end{array}
\]

\item For each inter-runnable variable $S$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\Tirv{\Ts{S}{I}}{V}
\end{array}
\]
where $V$ is the initial value attributed to $S$ if it exists, otherwise $V$ is $\Tvoid$.

\item For each data element $E$ of each required port $P$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\Tqelem{\Tep{E}{P}{I}}{N}{\epsilon}
\end{array}
\]
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued} (with capacity $N$); otherwise, a term
\[
\begin{array}{c}
	\Tdelem{\Tep{E}{P}{I}}{\Tfalse}{\text{nodata}}
\end{array}
\]

\item For each operation $O$ of each required port $P$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\Topres{\Tep{O}{P}{I}}{\epsilon}
\end{array}
\]

\item For each \emph{TimingEvent} triggering each runnable $R$ of each atomic component $I$, a term
\[
\begin{array}{c}
	\Ttimer{\Tr{R}{I}}{0.0}
\end{array}
\]
% where $T$ is the period of the event.
%TODO: Does this mean all timed runnables will want to act _immediately_ (at the same time)? No init delay?
